---
title: "sRNAs profiles Homo sapiens and SARS-CoV-2"
author: "Pascual Villalba"
date: "19/2/2021"
output: html_document
---

## 0. LOAD THE LIBRARIES AND OPTIONS

```{r}
library(dplyr)
options(dplyr.print_max = 1e9)
require(ggplot2)
library(ggsci)
library(ggExtra)
```

## 1. CREATE THE DIRECTORY WHERE THE PLOTS WILL BE SAVED

```{r}
if (!dir.exists("/path/to/results/04-library_profiles")){  # Creates the directory if it does not exist.
  dir.create("/path/to/results/04-library_profiles")
}
```

## 2. CREATE A LIST WITH THE ABSOLUTE PATH OF THE CSV FILES WITH ABSOLUTE AND RPM COUNTS

```{r}
## Absolute path of absolute counts.
exp_counts  = paste0(list.files("/path/to/results/01-library_absolute_counts", recursive = T, full.names = T))
## RPMs path of RPM counts.
exp_RPMs  = paste0(list.files("/path/to/results/02-library_RPM_counts", recursive = T, full.names = T))
```

## 3. CREATE THE TABLE DTFULL

Build a table containing the counts and frequence of the sequence lengths per library. This table will be saved as RPMs_and_Freq_by_ntds.tsv.

```{r}
## Number of iteration. It's necessary to build the global table.
counts= 0

for(i in 1:length(exp_counts)){
  ## Each iteration is a sample (library).
  ##i = 1
  file_counts = exp_counts[i]
  file_RPMs = exp_RPMs[i]
  
  ## Obtain the info: time (T1), sample (HSVd_1), group (HSVd) and n_group (1).
  path_file_list = unlist(strsplit(file_counts, "/"))
  time = path_file_list[length(path_file_list)-2]
  name_file = path_file_list[length(path_file_list)]
  sample = unlist(strsplit(name_file, ".", fixed = TRUE))[1]
  group = unlist(strsplit(sample, "-"))[1]
  n_group = unlist(strsplit(sample, "-"))[2]
  
  ## Load the absolute and RPM counts table.
  counts_table = read.csv(file_counts, sep=",", header = T, stringsAsFactors = F)
  RPMs_table = read.csv(file_RPMs, sep=",", header = T, stringsAsFactors = F)
  
  ## Create a empty table with the counts, RPMs and percentage by the number of nucleotides for each sample.
  dt = data.frame(nt=c(12:34), stringsAsFactors = F)
  dt$counts = 0
  dt$total = 0
  dt$RPM = 0
  dt$percentage = 0
  dt$sample = sample
  dt$group = group
  dt$n_group = n_group
  dt$time = time
  
  ## Group the absolute and RPM counts of the absolute and RPM counts table by number of nucleotides.
  for(j in 1:dim(dt)[1]){
    # j=1
    subset_counts = counts_table[nchar(counts_table$seq) == dt$nt[j],] # Filtrar la tabla por aquellas secuencias de longitud 12-34.
    dt$counts[j] = sum(subset_counts$counts)
    subset_RPMs = RPMs_table[nchar(RPMs_table$seq) == dt$nt[j],] # Filtrar la tabla por aquellas secuencias de longitud 12-34.
    dt$RPM[j] = sum(subset_RPMs$RPMs)
  }
  
  ## Calculate the library size.
  sumtotal = sum(dt$counts)
  dt$total = sumtotal
  
  ## Calculate the percentage by number of nucleotides.
  dt$percentage = (dt$counts/dt$total)*100
  
  ## Build the global table. Add the table created in each iteration.
  if(counts == 0){
    dtfull = dt
    counts = counts + 1
  }else{
    dtfull = rbind(dtfull, dt)
  }
}

## Save the table.
write.table(dtfull,"/path/to/results/04-library_profiles/RPMs_and_Freq_by_ntds.tsv", 
            sep="\t", dec=",", row.names = F, quote = F)

## Show dtfull
dtfull

## Remove the data and values of the R environment.
#rm(list = c('counts_table', 'RPMs_table', 'dt', 'i', 'j', 'name_file', 'exp_counts', 'exp_RPMs', 'file_counts', 'file_RPMs', 'group', 'n_group', 'path_file_list', 'sample', 'sumtotal', 'time', 'subset_counts', 'subset_RPMs', 'counts'))
```



## 4. CREATE THE TABLE DTFULL_FINAL

The same info summarized by group. 
```{r}
## This is the function for calculating the standard deviation.
se = function(x) sqrt(var(x)/length(x)) 

## Group the previous table by group (HSVd), time (T1) and nt (20)
dtfull_FINAL = dtfull %>% 
  group_by(group, time, nt) %>%
  summarise(
    mean.RPM = mean(RPM), ## Mean RPMs.
    se.RPM = se(RPM), ## Standard deviation of RPMs.
    mean.Relative = mean(percentage), ## Mean of percentage.
    se.Relative = se(percentage)) ## Standard deviation of percentage. 

## Save the table.
write.table(dtfull_FINAL,"/path/to/results/04-library_profiles/RPMs_and_Freq_by_ntds-FINAL.tsv",
            sep="\t", dec=",", row.names = F, quote = F)

## Show dtfull_FINAL
dtfull_FINAL
```


## 5. PLOTS


```{r}
## Convert the group to factor.
dtfull_FINAL$group <- ifelse(dtfull_FINAL$group == "acute", "Severe", dtfull_FINAL$group)
dtfull_FINAL$group <- ifelse(dtfull_FINAL$group == "mild", "Moderate", dtfull_FINAL$group)
dtfull_FINAL$group <- ifelse(dtfull_FINAL$group == "control", "Control", dtfull_FINAL$group)
dtfull_FINAL$group = factor(dtfull_FINAL$group, levels=c("Control","Severe","Moderate"))
```


### 5.1 PLOT WITH A FACET_WRAP BY TIME

```{r}
options(scipen = 10)
time_names <-  c("Control", "T1", "T2")
names(time_names) <- c("T0", "T1", "T2")

ggplot(dtfull_FINAL, aes(x=nt, y=mean.RPM, ymin=mean.RPM-se.RPM, ymax=mean.RPM+se.RPM, fill=group)) +
  geom_bar(position=position_dodge(), aes(y=mean.RPM), stat="identity") +
  geom_errorbar (position=position_dodge(width=0.9), colour="black")+
  scale_x_discrete(breaks=c(12:34),limits=12:34)+
  scale_fill_lancet(alpha = 0.7)+
  theme_bw()+removeGrid( x = T, y = T  )+ xlab("Read size (nt)")+ ylab("RPM")+
  #geom_segment(aes(x = 19, y = 0, xend = 26, yend = 0), linetype=2, colour = "darkgray", size = 0.4 )+
  facet_wrap( ~ time, ncol=3, labeller = as_labeller(time_names))+
  theme(legend.position="bottom", axis.text.x = element_text(angle=65, vjust=0.6), plot.title = element_text(hjust = 0.5)) +
  theme(legend.title = element_blank())

## Save the plot.
ggsave("/path/to/results/04-library_profiles/ALL-sRNAs_profile_RPMs_1_hor.png", height = 3.25, width = 9, dpi = 300)
```






