---
title: "tRNA analysis"
author: "Julia Corell Sierra"
date: "2023-06-12"
output: html_document
---


In this script you'll find all the code necessary to produce the figures and tables related to the tRNA analysis.
```{r}
library(tidyverse)
library(DESeq2)
library(ggbreak)
```

```{r}
padj.cutoff <- 0.05
anno_file <- read.table("/path/to/annotation_results/unitas_simplified_annnotation/curated_unitas_withtiny_annotation.tsv", sep="\t", header=TRUE)
RPMs_table <- read.table("/path/to/results/03-Fusion_tables/Filtered_RPMs-outer_filt5.tsv", sep="\t", header = TRUE)
sampletable <- read.table("../../Additional_data/samplesheet.tsv", header=T, sep="\t")
DE_T1_acute <- read.table("/path/to/DEA_results/results_DE_T1acute.tsv", sep="\t", header=TRUE)
DE_T1_acute_sig <- DE_T1_acute %>%
  dplyr::filter(padj < padj.cutoff)
DE_T1_mild <- read.table("/path/to/DEA_results/results_DE_T1mild.tsv", sep="\t", header=TRUE)
DE_T1_mild_sig <- DE_T1_mild %>%
  dplyr::filter(padj < padj.cutoff)
DE_T2_acute <- read.table("/path/to/DEA_results/results_DE_T2acute.tsv", sep="\t", header=TRUE)
DE_T2_acute_sig <- DE_T2_acute %>%
  dplyr::filter(padj < padj.cutoff)
DE_T2_mild <- read.table("/path/to/DEA_results/results_DE_T2mild.tsv", sep="\t", header=TRUE)
DE_T2_mild_sig <- DE_T2_mild %>%
  dplyr::filter(padj < padj.cutoff)

DE_T1_acute_sig_ann <- merge(DE_T1_acute_sig, anno_file, by.x = "seq", by.y = "Seq", all.x = TRUE)
DE_T1_mild_sig_ann <-  merge(DE_T1_mild_sig, anno_file, by.x = "seq", by.y = "Seq", all.x = TRUE)
DE_T2_acute_sig_ann <- merge(DE_T2_acute_sig, anno_file, by.x = "seq", by.y = "Seq", all.x = TRUE)
DE_T2_mild_sig_ann <-  merge(DE_T2_mild_sig, anno_file, by.x = "seq", by.y = "Seq", all.x = TRUE)
```


```{r}
DE_T1_acute_tRNA <- DE_T1_acute_sig_ann %>%
  dplyr::filter(grepl("tRNA", DE_T1_acute_sig_ann$RNA_class))

DE_T1_mild_tRNA <- DE_T1_mild_sig_ann %>%
  dplyr::filter(grepl("tRNA", DE_T1_mild_sig_ann$RNA_class))

DE_T2_acute_tRNA <- DE_T2_acute_sig_ann %>%
  dplyr::filter(grepl("tRNA", DE_T2_acute_sig_ann$RNA_class))

DE_T2_mild_tRNA <- DE_T2_mild_sig_ann %>%
  dplyr::filter(grepl("tRNA", DE_T2_mild_sig_ann$RNA_class))

#DE_tRNAs <- Reduce (union, list(DE_T1_acute_tRNA$seq, DE_T1_mild_tRNA$seq, DE_T2_acute_tRNA$seq, DE_T2_mild_tRNA$seq))
```


***

#### LFC Boxplots supplemental figure

Let's take a look at the tRNA sequences behavior. Supplemental Figure with the LFC of every tRNA.

```{r}
DE_T1_acute_tRNA$ID_DB_cut <- str_replace(DE_T1_acute_tRNA$RNA_ID, "-[A-Z]{3}-[0-9][0-9]?-[0-9][0-9]?", "")
DE_T1_acute_tRNA$ID_DB_cut <- str_replace(DE_T1_acute_tRNA$ID_DB_cut, "tRNA-", "")
DE_T1_acute_tRNA <- DE_T1_acute_tRNA %>%
  filter(!grepl("MT", DE_T1_acute_tRNA$ID_DB_cut))
DE_T1_acute_tRNA <- DE_T1_acute_tRNA %>%
  filter(!grepl("tRF-1|misc-tRF", DE_T1_acute_tRNA$tRNA_class))

DE_T2_acute_tRNA$ID_DB_cut <- str_replace(DE_T2_acute_tRNA$RNA_ID, "-[A-Z]{3}-[0-9][0-9]?-[0-9][0-9]?", "")
DE_T2_acute_tRNA$ID_DB_cut <- str_replace(DE_T2_acute_tRNA$ID_DB_cut, "tRNA-", "")
DE_T2_acute_tRNA <- DE_T2_acute_tRNA %>%
  filter(!grepl("MT", DE_T2_acute_tRNA$ID_DB_cut))
DE_T2_acute_tRNA <- DE_T2_acute_tRNA %>%
  filter(!grepl("tRF-1|misc-tRF", DE_T2_acute_tRNA$tRNA_class))

DE_T1_mild_tRNA$ID_DB_cut <- str_replace(DE_T1_mild_tRNA$RNA_ID, "-[A-Z]{3}-[0-9][0-9]?-[0-9][0-9]?", "")
DE_T1_mild_tRNA$ID_DB_cut <- str_replace(DE_T1_mild_tRNA$ID_DB_cut, "tRNA-", "")
DE_T1_mild_tRNA <- DE_T1_mild_tRNA %>%
  filter(!grepl("MT", DE_T1_mild_tRNA$ID_DB_cut))
DE_T1_mild_tRNA <- DE_T1_mild_tRNA %>%
  filter(!grepl("tRF-1|misc-tRF", DE_T1_mild_tRNA$tRNA_class))

DE_T2_mild_tRNA$ID_DB_cut <- str_replace(DE_T2_mild_tRNA$RNA_ID, "-[A-Z]{3}-[0-9][0-9]?-[0-9][0-9]?", "")
DE_T2_mild_tRNA$ID_DB_cut <- str_replace(DE_T2_mild_tRNA$ID_DB_cut, "tRNA-", "")
DE_T2_mild_tRNA <- DE_T2_mild_tRNA %>%
  filter(!grepl("MT", DE_T2_mild_tRNA$ID_DB_cut))
DE_T2_mild_tRNA <- DE_T2_mild_tRNA %>%
  filter(!grepl("tRF-1|misc-tRF", DE_T2_mild_tRNA$tRNA_class))

aa_list <- Reduce(union, list(DE_T1_acute_tRNA$ID_DB_cut, DE_T1_mild_tRNA$ID_DB_cut, DE_T2_acute_tRNA$ID_DB_cut, DE_T2_mild_tRNA$ID_DB_cut))
aa_list <- sort(aa_list)
```

```{r}
DE_T1_acute_tRNA$tRNA_class <- ifelse(grepl("CCA", DE_T1_acute_tRNA$tRNA_class), "3p-tRF", DE_T1_acute_tRNA$tRNA_class)

DE_T1_acute_tRNA$ID_DB_cut <- factor(DE_T1_acute_tRNA$ID_DB_cut, levels= aa_list)

##Alternative plot without misc-tRF and tRF-1

##

ggplot(DE_T1_acute_tRNA, aes(x=ID_DB_cut, y=lfcShrunk)) +
  geom_boxplot(outlier.shape=NA, fill="#da6a65", alpha=0.7) +
  geom_jitter(size=2, alpha=0.7, aes(shape= tRNA_class)) +
  scale_shape_manual("tsRNA class", values=c(17,19,2,1,8,5)) +
  scale_x_discrete(drop=FALSE) +
  theme_bw() +
  geom_hline(yintercept=0, color="black") +
  theme(#legend.position = "none",
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(color="black", vjust=0.5),
    axis.text.y = element_text(color="black")) + ggtitle("Severe (T1) tsRNAs expression") + ylab("Log2 fold change") +
  xlab("tRNA")

ggsave("/path/to/tRNA_results/T1_acute_tRNAs_aa_filt.png", height=5, width=15, dpi=600)
```
```{r}
summary_table_tRNA_T1acute <- data.frame(tRNA=rep(unique(DE_T1_acute_tRNA$ID_DB_cut), each=length(unique(DE_T1_acute_tRNA$tRNA_class))), 
                                         tRNA_type=rep(unique(DE_T1_acute_tRNA$tRNA_class), length(unique(DE_T1_acute_tRNA$ID_DB_cut))), 
                                                       Up="NA", Down="NA")
for (i in 1:nrow(summary_table_tRNA_T1acute)){
  summary_table_tRNA_T1acute$Up[i] <- nrow(DE_T1_acute_tRNA[DE_T1_acute_tRNA$ID_DB_cut==summary_table_tRNA_T1acute$tRNA[i] & 
                                                              DE_T1_acute_tRNA$tRNA_class==summary_table_tRNA_T1acute$tRNA_type[i] & 
                                                              DE_T1_acute_tRNA$lfcShrunk > 0,])
  summary_table_tRNA_T1acute$Down[i] <- nrow(DE_T1_acute_tRNA[DE_T1_acute_tRNA$ID_DB_cut==summary_table_tRNA_T1acute$tRNA[i] & 
                                                              DE_T1_acute_tRNA$tRNA_class==summary_table_tRNA_T1acute$tRNA_type[i] & 
                                                              DE_T1_acute_tRNA$lfcShrunk < 0,])
}
write.table(summary_table_tRNA_T1acute, "/path/to/tRNA_results/summary_table_tRNA_T1_acute.tsv", sep="\t", row.names=FALSE)
```

```{r}
DE_T2_acute_tRNA$tRNA_class <- ifelse(grepl("CCA", DE_T2_acute_tRNA$tRNA_class), "3p-tRF", DE_T2_acute_tRNA$tRNA_class)

DE_T2_acute_tRNA$ID_DB_cut <- factor(DE_T2_acute_tRNA$ID_DB_cut, levels= aa_list)

##Alternative plot without misc-tRF and tRF-1
##

ggplot(DE_T2_acute_tRNA, aes(x=ID_DB_cut, y=lfcShrunk)) +
  geom_boxplot(outlier.shape=NA, fill="#ED0000B2", alpha=0.7) +
  geom_jitter(size=2, alpha=0.7, aes(shape= tRNA_class)) +
  scale_shape_manual("tsRNA class", values=c(19,2,1,8,5)) +
  scale_x_discrete(drop=FALSE) +
  theme_bw() +
  geom_hline(yintercept=0, color="black") +
  theme(#legend.position = "none",
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(color="black", vjust = 0.5),
    axis.text.y = element_text(color="black")) + ggtitle("Severe (T2) tsRNAs expression") + ylab("Log2 fold change") +
  xlab("tRNA")

ggsave("/path/to/tRNA_results/T2_acute_tRNAs_aa_exp_filt.png", height=5, width=15, dpi=600)
```
```{r}
summary_table_tRNA_T2acute <- data.frame(tRNA=rep(unique(DE_T2_acute_tRNA$ID_DB_cut), each=length(unique(DE_T1_acute_tRNA$tRNA_class))), 
                                         tRNA_type=rep(unique(DE_T1_acute_tRNA$tRNA_class), length(unique(DE_T2_acute_tRNA$ID_DB_cut))), 
                                                       Up="NA", Down="NA")
for (i in 1:nrow(summary_table_tRNA_T2acute)){
  summary_table_tRNA_T2acute$Up[i] <- nrow(DE_T2_acute_tRNA[DE_T2_acute_tRNA$ID_DB_cut==summary_table_tRNA_T2acute$tRNA[i] & 
                                                              DE_T2_acute_tRNA$tRNA_class==summary_table_tRNA_T2acute$tRNA_type[i] & 
                                                              DE_T2_acute_tRNA$lfcShrunk > 0,])
  summary_table_tRNA_T2acute$Down[i] <- nrow(DE_T2_acute_tRNA[DE_T2_acute_tRNA$ID_DB_cut==summary_table_tRNA_T2acute$tRNA[i] & 
                                                              DE_T2_acute_tRNA$tRNA_class==summary_table_tRNA_T2acute$tRNA_type[i] & 
                                                              DE_T2_acute_tRNA$lfcShrunk < 0,])
}
write.table(summary_table_tRNA_T2acute, "/path/to/tRNA_results/summary_table_tRNA_T2_acute.tsv", sep="\t", row.names=FALSE)
```

```{r}
DE_T1_mild_tRNA$tRNA_class <- ifelse(grepl("CCA", DE_T1_mild_tRNA$tRNA_class), "3p-tRF", DE_T1_mild_tRNA$tRNA_class)

DE_T1_mild_tRNA$ID_DB_cut <- factor(DE_T1_mild_tRNA$ID_DB_cut, levels= aa_list)

##Alternative plot without misc-tRF and tRF-1

##

ggplot(DE_T1_mild_tRNA, aes(x=ID_DB_cut, y=lfcShrunk)) +
  geom_boxplot(outlier.shape=NA, fill="lightgreen", alpha=0.7) +
  geom_jitter(size=2, alpha=0.7, aes(shape= tRNA_class)) +
  scale_shape_manual("tsRNA class", values=c(19,2,1,8)) +
  scale_x_discrete(drop=FALSE) +
  theme_bw() +
  geom_hline(yintercept=0, color="black") +
  theme(#legend.position = "none",
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(color="black", hjust = 1, vjust = 0.5),
    axis.text.y = element_text(color="black")) + ggtitle("Moderate (T1) tsRNAs expression") + ylab("Log2 fold change") +
  xlab("tRNA")

ggsave("/path/to/tRNA_results/T1_mild_tRNAs_aa_exp_filt.png", height=5, width=15, dpi=600)
```
```{r}
summary_table_tRNA_T1mild <- data.frame(tRNA=rep(unique(DE_T1_mild_tRNA$ID_DB_cut), each=length(unique(DE_T1_acute_tRNA$tRNA_class))), 
                                         tRNA_type=rep(unique(DE_T1_acute_tRNA$tRNA_class), length(unique(DE_T1_mild_tRNA$ID_DB_cut))), 
                                                       Up="NA", Down="NA")
for (i in 1:nrow(summary_table_tRNA_T1mild)){
  summary_table_tRNA_T1mild$Up[i] <- nrow(DE_T1_mild_tRNA[DE_T1_mild_tRNA$ID_DB_cut==summary_table_tRNA_T1mild$tRNA[i] & 
                                                              DE_T1_mild_tRNA$tRNA_class==summary_table_tRNA_T1mild$tRNA_type[i] & 
                                                              DE_T1_mild_tRNA$lfcShrunk > 0,])
  summary_table_tRNA_T1mild$Down[i] <- nrow(DE_T1_mild_tRNA[DE_T1_mild_tRNA$ID_DB_cut==summary_table_tRNA_T1mild$tRNA[i] & 
                                                              DE_T1_mild_tRNA$tRNA_class==summary_table_tRNA_T1mild$tRNA_type[i] & 
                                                              DE_T1_mild_tRNA$lfcShrunk < 0,])
}
write.table(summary_table_tRNA_T1mild, "/path/to/tRNA_results/summary_table_tRNA_T1_mild.tsv", sep="\t", row.names=FALSE)
```

```{r}
DE_T2_mild_tRNA$tRNA_class <- ifelse(grepl("CCA", DE_T2_mild_tRNA$tRNA_class), "3p-tRF", DE_T2_mild_tRNA$tRNA_class)

DE_T2_mild_tRNA$ID_DB_cut <- factor(DE_T2_mild_tRNA$ID_DB_cut, levels= aa_list)

##Alternative plot without misc-tRF and tRF-1

##

ggplot(DE_T2_mild_tRNA, aes(x=ID_DB_cut, y=lfcShrunk)) +
  geom_boxplot(outlier.shape=NA, fill="darkgreen", alpha=0.7) +
  geom_jitter(size=2, alpha=0.7, aes(shape= tRNA_class)) +
  scale_shape_manual("tsRNA class", values=c(19,2,1,8)) +
  scale_x_discrete(drop=FALSE) +
  theme_bw() +
  geom_hline(yintercept=0, color="black") +
  theme(#legend.position = "none",
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(color="black",  vjust = 0.5),
    axis.text.y = element_text(color="black")) + ggtitle("Moderate (T2) tsRNAs expression") + ylab("Log2 fold change") +
  xlab("tRNA")

ggsave("/path/to/tRNA_results/T2_mild_tRNAs_aa_exp_filt.png", height=5, width=15, dpi=600)
```
```{r}
summary_table_tRNA_T2mild <- data.frame(tRNA=rep(unique(DE_T2_mild_tRNA$ID_DB_cut), each=length(unique(DE_T1_acute_tRNA$tRNA_class))), 
                                         tRNA_type=rep(unique(DE_T1_acute_tRNA$tRNA_class), length(unique(DE_T2_mild_tRNA$ID_DB_cut))), 
                                                       Up="NA", Down="NA")
for (i in 1:nrow(summary_table_tRNA_T2mild)){
  summary_table_tRNA_T2mild$Up[i] <- nrow(DE_T2_mild_tRNA[DE_T2_mild_tRNA$ID_DB_cut==summary_table_tRNA_T2mild$tRNA[i] & 
                                                              DE_T2_mild_tRNA$tRNA_class==summary_table_tRNA_T2mild$tRNA_type[i] & 
                                                              DE_T2_mild_tRNA$lfcShrunk > 0,])
  summary_table_tRNA_T2mild$Down[i] <- nrow(DE_T2_mild_tRNA[DE_T2_mild_tRNA$ID_DB_cut==summary_table_tRNA_T2mild$tRNA[i] & 
                                                              DE_T2_mild_tRNA$tRNA_class==summary_table_tRNA_T2mild$tRNA_type[i] & 
                                                              DE_T2_mild_tRNA$lfcShrunk < 0,])
}
write.table(summary_table_tRNA_T2mild, "/path/to/tRNA_results/summary_table_tRNA_T2_mild.tsv", sep="\t", row.names=FALSE)
```


### Supplemental table

First we need a summary of the tRNA annotation of our DE sequences.
```{r}
DE_T1_acute_sel <- DE_T1_acute_tRNA[, c("seq", "RNA_ID", "tRNA_class", "ID_DB_cut")]
DE_T2_acute_sel <- DE_T2_acute_tRNA[, c("seq", "RNA_ID", "tRNA_class", "ID_DB_cut")]
DE_T1_mild_sel <- DE_T1_mild_tRNA[, c("seq", "RNA_ID", "tRNA_class", "ID_DB_cut")]
DE_T2_mild_sel <- DE_T2_mild_tRNA[, c("seq", "RNA_ID", "tRNA_class", "ID_DB_cut")]

summ_tRNA_anno <- rbind(DE_T1_acute_sel, DE_T2_acute_sel, DE_T1_mild_sel, DE_T2_mild_sel)
summ_tRNA_anno <- unique(summ_tRNA_anno)
```

```{r}
summary_tRNA_table <- data.frame(seq=summ_tRNA_anno$seq, Unitas_ID=summ_tRNA_anno$RNA_ID, Unitas_tRNA_class=summ_tRNA_anno$tRNA_class,
                                 Isotype=summ_tRNA_anno$ID_DB_cut, Severe_T1_LFC=NA, Severe_T1_padj=NA, Severe_T2_LFC=NA, Severe_T2_padj=NA, 
                                 Moderate_T1_LFC=NA, Moderate_T1_padj=NA,Moderate_T2_LFC=NA, Moderate_T2_padj=NA)

for (tRNA_seq in summ_tRNA_anno$seq) {
  summary_tRNA_table$Severe_T1_LFC[summary_tRNA_table$seq == tRNA_seq] <- ifelse(tRNA_seq %in% DE_T1_acute_tRNA$seq, DE_T1_acute_tRNA[DE_T1_acute_tRNA$seq == tRNA_seq, "lfcShrunk"], NA)
  summary_tRNA_table$Severe_T1_padj[summary_tRNA_table$seq == tRNA_seq] <- ifelse(tRNA_seq %in% DE_T1_acute_tRNA$seq, DE_T1_acute_tRNA[DE_T1_acute_tRNA$seq == tRNA_seq, "padj"], NA)
  summary_tRNA_table$Severe_T2_LFC[summary_tRNA_table$seq == tRNA_seq] <- ifelse(tRNA_seq %in% DE_T2_acute_tRNA$seq, DE_T2_acute_tRNA[DE_T2_acute_tRNA$seq == tRNA_seq, "lfcShrunk"], NA)
  summary_tRNA_table$Severe_T2_padj[summary_tRNA_table$seq == tRNA_seq] <- ifelse(tRNA_seq %in% DE_T2_acute_tRNA$seq, DE_T1_acute_tRNA[DE_T2_acute_tRNA$seq == tRNA_seq, "padj"], NA)
  summary_tRNA_table$Moderate_T1_LFC[summary_tRNA_table$seq == tRNA_seq] <- ifelse(tRNA_seq %in% DE_T1_mild_tRNA$seq, DE_T1_mild_tRNA[DE_T1_mild_tRNA$seq == tRNA_seq, "lfcShrunk"], NA)
  summary_tRNA_table$Moderate_T1_padj[summary_tRNA_table$seq == tRNA_seq] <- ifelse(tRNA_seq %in% DE_T1_mild_tRNA$seq, DE_T1_mild_tRNA[DE_T1_mild_tRNA$seq == tRNA_seq, "padj"], NA)
  summary_tRNA_table$Moderate_T2_LFC[summary_tRNA_table$seq == tRNA_seq] <- ifelse(tRNA_seq %in% DE_T2_mild_tRNA$seq, DE_T2_mild_tRNA[DE_T2_mild_tRNA$seq == tRNA_seq, "lfcShrunk"], NA)
  summary_tRNA_table$Moderate_T2_padj[summary_tRNA_table$seq == tRNA_seq] <- ifelse(tRNA_seq %in% DE_T2_mild_tRNA$seq, DE_T1_mild_tRNA[DE_T2_mild_tRNA$seq == tRNA_seq, "padj"], NA)
}

summary_tRNA_table <- summary_tRNA_table[order(summary_tRNA_table$Isotype),]
write.table(summary_tRNA_table, "/path/to/tRNA_results/Supp_tRNA.tsv", sep="\t", row.names=FALSE)
```

### Accumulation plots of Figure 4

Let's study in more detail what are the accumulation profiles of the different tsRNAs from tRNA Glu and tRNA Gly (the DE sequences which accumulate greater).

```{r}
#Create ID_DB_cut and unique category for 3p-tRF
DE_T1_acute_tRNA$ID_DB_cut <- str_replace(DE_T1_acute_tRNA$RNA_ID, "-[A-Z]{3}-[0-9][0-9]?-[0-9][0-9]?", "")
DE_T1_acute_tRNA$tRNA_class <- ifelse(grepl("CCA", DE_T1_acute_tRNA$tRNA_class), "3p-tRF", DE_T1_acute_tRNA$tRNA_class)
DE_T2_acute_tRNA$ID_DB_cut <- str_replace(DE_T2_acute_tRNA$RNA_ID, "-[A-Z]{3}-[0-9][0-9]?-[0-9][0-9]?", "")
DE_T2_acute_tRNA$tRNA_class <- ifelse(grepl("CCA", DE_T2_acute_tRNA$tRNA_class), "3p-tRF", DE_T2_acute_tRNA$tRNA_class)
DE_T1_mild_tRNA$ID_DB_cut <- str_replace(DE_T1_mild_tRNA$RNA_ID, "-[A-Z]{3}-[0-9][0-9]?-[0-9][0-9]?", "")
DE_T1_mild_tRNA$tRNA_class <- ifelse(grepl("CCA", DE_T1_mild_tRNA$tRNA_class), "3p-tRF", DE_T1_mild_tRNA$tRNA_class)
DE_T2_mild_tRNA$ID_DB_cut <- str_replace(DE_T2_mild_tRNA$RNA_ID, "-[A-Z]{3}-[0-9][0-9]?-[0-9][0-9]?", "")
DE_T2_mild_tRNA$tRNA_class <- ifelse(grepl("CCA", DE_T2_mild_tRNA$tRNA_class), "3p-tRF", DE_T2_mild_tRNA$tRNA_class)
```


### Glu and Gly

#### 5p-tR-half
This chunk, that will be repeated for every tsRNA type, creates a table with the sum of the accumulation of each of the forms of Glu and Gly 5p-tR-half. 
```{r}
sample_RPM_GluGly_5p_tRhalf <- data.frame(Sample=sampletable$SampleName[-c(1:8)], Group=sampletable$Group[-c(1:8)], Glu_control=NA, 
                                          Glu_infected=NA, Gly_control=NA, Gly_infected=NA)

for (case in c("T1_acute", "T1_mild", "T2_acute", "T2_mild")){
  for (aa in c("Glu", "Gly")){
    df <- get(paste0("DE_", case, "_tRNA"))
    df_subset <- df[df$ID_DB_cut == paste0("tRNA-",aa) & df$tRNA_class == "5p-tR-half",]
    seqs <- df_subset$seq
    col_control <- colnames(RPMs_table)[which(grepl("control", colnames(RPMs_table)))]
    col_case <- colnames(RPMs_table)[which(grepl(case, colnames(RPMs_table)))]
    df_RPM <- RPMs_table[RPMs_table$seq %in% seqs, c("seq",col_control,col_case)]
    df_RPM <- df_RPM %>% #Sum every seq
      bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"Total")))
    sample_RPM_GluGly_5p_tRhalf[sample_RPM_GluGly_5p_tRhalf$Group == case, paste0(aa, "_control")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl("control", colnames(df_RPM))])
    sample_RPM_GluGly_5p_tRhalf[sample_RPM_GluGly_5p_tRhalf$Group == case, paste0(aa, "_infected")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl(case, colnames(df_RPM))])
  }
}

sample_RPM_GluGly_5p_tRhalf_long <- sample_RPM_GluGly_5p_tRhalf%>%
  pivot_longer(c(3:6),
               names_to="Condition",
               values_to="RPM")
sample_RPM_GluGly_5p_tRhalf_long$Color <- ifelse(grepl("control", sample_RPM_GluGly_5p_tRhalf_long$Condition), "Control", paste0(sample_RPM_GluGly_5p_tRhalf_long$Group,"_Infected"))
```

```{r}
sample_RPM_GluGly_5p_tRhalf_long$Group <- factor(sample_RPM_GluGly_5p_tRhalf_long$Group, levels = c("T1_acute", "T2_acute", "T1_mild", "T2_mild"))
state.labs <- c("Severe (T1)", "Severe (T2)", "Moderate (T1)", "Moderate (T2)")
names(state.labs) <- c("T1_acute", "T2_acute", "T1_mild", "T2_mild")
sample_RPM_GluGly_5p_tRhalf_long$Label <- gsub("T[012]\\_(acute|mild|control)-", "", sample_RPM_GluGly_5p_tRhalf_long$Sample)

ggplot(sample_RPM_GluGly_5p_tRhalf_long, aes(x=Condition, y=RPM, fill=Color, color=Color)) +
  geom_violin(alpha=0.7) +
  geom_boxplot(width=0.25, outlier.shape = NA, alpha=0.7, color="black") +
  geom_jitter(position=position_jitter(0.2),  alpha=0.7) +
  scale_fill_manual(values=c("#00468BB2", "#da6a65", "lightgreen", "#ED0000B2", "darkgreen")) +
  scale_color_manual(values=c("#00468BB2", "#da6a65", "lightgreen","#ED0000B2", "darkgreen")) +
  #geom_text_repel(max.overlaps = 50) +
  theme_classic() +
  facet_wrap(~Group, nrow = 1, strip.position = "bottom", labeller = labeller(Group=state.labs)) +
  theme(legend.position="none",
        axis.title = element_text(size=15),
        plot.title = element_text(size = rel(2)),
        axis.text.x = element_text(angle=45, vjust=1, hjust=1, color="black", size=12),
        axis.text.y = element_text(color="black", size=12),
        panel.grid = element_blank(),
        strip.background = element_rect(fill="#ebeff1"),
        strip.text = element_text(size=12),
        panel.border = element_blank(),
        axis.line.y = element_line(color="black"),
        axis.line.x = element_blank()) + ggtitle("5' tR-half") + xlab("")
ggsave("/path/to/tRNA_results/GluGly_violin_5p_tRhalf_withlabels.png", height = 5, width = 7)
```

#### 3p-tR-half

```{r}
sample_RPM_GluGly_3p_tRhalf <- data.frame(Sample=sampletable$SampleName[-c(1:8)], Group=sampletable$Group[-c(1:8)], Glu_control=NA, 
                                          Glu_infected=NA, Gly_control=NA, Gly_infected=NA)

for (case in c("T1_acute", "T1_mild", "T2_acute", "T2_mild")){
  for (aa in c("Glu", "Gly")){
    df <- get(paste0("DE_", case, "_tRNA"))
    df_subset <- df[df$ID_DB_cut == paste0("tRNA-",aa) & df$tRNA_class == "3p-tR-half",]
    seqs <- df_subset$seq
    col_control <- colnames(RPMs_table)[which(grepl("control", colnames(RPMs_table)))]
    col_case <- colnames(RPMs_table)[which(grepl(case, colnames(RPMs_table)))]
    df_RPM <- RPMs_table[RPMs_table$seq %in% seqs, c("seq",col_control,col_case)]
    df_RPM <- df_RPM %>% #Sum every seq
      bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"Total")))
    sample_RPM_GluGly_3p_tRhalf[sample_RPM_GluGly_3p_tRhalf$Group == case, paste0(aa, "_control")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl("control", colnames(df_RPM))])
    sample_RPM_GluGly_3p_tRhalf[sample_RPM_GluGly_3p_tRhalf$Group == case, paste0(aa, "_infected")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl(case, colnames(df_RPM))])
  }
}

sample_RPM_GluGly_3p_tRhalf_long <- sample_RPM_GluGly_3p_tRhalf%>%
  pivot_longer(c(3:6),
               names_to="Condition",
               values_to="RPM")
sample_RPM_GluGly_3p_tRhalf_long$Color <- ifelse(grepl("control", sample_RPM_GluGly_3p_tRhalf_long$Condition), "Control", paste0(sample_RPM_GluGly_3p_tRhalf_long$Group,"_Infected"))
```


```{r}
sample_RPM_GluGly_3p_tRhalf_long$Group <- factor(sample_RPM_GluGly_3p_tRhalf_long$Group, levels = c("T1_acute", "T2_acute", "T1_mild", "T2_mild"))
state.labs <- c("Severe (T1)", "Severe (T2)", "Moderate (T1)", "Moderate (T2)")
names(state.labs) <- c("T1_acute", "T2_acute", "T1_mild", "T2_mild")
sample_RPM_GluGly_3p_tRhalf_long$Label <- gsub("T[012]\\_(acute|mild|control)-", "", sample_RPM_GluGly_3p_tRhalf_long$Sample)

ggplot(sample_RPM_GluGly_3p_tRhalf_long, aes(x=Condition, y=RPM, fill=Color, color=Color)) +
  geom_violin(alpha=0.7) +
  geom_boxplot(width=0.25, outlier.shape = NA, alpha=0.7, color="black") +
  geom_jitter(position=position_jitter(0.2), alpha=0.7) +
  scale_fill_manual(values=c("#00468BB2", "#da6a65", "lightgreen", "#ED0000B2", "darkgreen")) +
  scale_color_manual(values=c("#00468BB2", "#da6a65", "lightgreen","#ED0000B2", "darkgreen")) +
  #geom_text_repel() +
  theme_classic() +
  facet_wrap(~Group, nrow = 1, strip.position = "bottom", labeller = labeller(Group=state.labs)) +
  theme(legend.position="none",
        axis.title = element_text(size=15),
        plot.title = element_text(size = rel(2)),
        axis.text.x = element_text(angle=45, vjust=1, hjust=1, color="black", size=12),
        axis.text.y = element_text(color="black", size=12),
        panel.grid = element_blank(),
        strip.background = element_rect(fill="#ebeff1"),
        strip.text = element_text(size=12),
        panel.border = element_blank(),
        axis.line.y = element_line(color="black"),
        axis.line.x = element_blank()) + ggtitle("3' tR-half") + xlab("") + ylim(c(-1,100))
ggsave("/path/to/tRNA_results/GluGly_violin_3p_tRhalf_withlabels.png", height = 5, width = 7)
```

#### 5p-tRF

```{r}
sample_RPM_GluGly_5p_tRF <- data.frame(Sample=sampletable$SampleName[-c(1:8)], Group=sampletable$Group[-c(1:8)], Glu_control=NA, 
                                          Glu_infected=NA, Gly_control=NA, Gly_infected=NA)

for (case in c("T1_acute", "T1_mild", "T2_acute", "T2_mild")){
  for (aa in c("Glu", "Gly")){
    df <- get(paste0("DE_", case, "_tRNA"))
    df_subset <- df[df$ID_DB_cut == paste0("tRNA-",aa) & df$tRNA_class == "5p-tRF",]
    seqs <- df_subset$seq
    col_control <- colnames(RPMs_table)[which(grepl("control", colnames(RPMs_table)))]
    col_case <- colnames(RPMs_table)[which(grepl(case, colnames(RPMs_table)))]
    df_RPM <- RPMs_table[RPMs_table$seq %in% seqs, c("seq",col_control,col_case)]
    df_RPM <- df_RPM %>% #Sum every seq
      bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"Total")))
    sample_RPM_GluGly_5p_tRF[sample_RPM_GluGly_5p_tRF$Group == case, paste0(aa, "_control")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl("control", colnames(df_RPM))])
    sample_RPM_GluGly_5p_tRF[sample_RPM_GluGly_5p_tRF$Group == case, paste0(aa, "_infected")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl(case, colnames(df_RPM))])
  }
}

sample_RPM_GluGly_5p_tRF_long <- sample_RPM_GluGly_5p_tRF%>%
  pivot_longer(c(3:6),
               names_to="Condition",
               values_to="RPM")
sample_RPM_GluGly_5p_tRF_long$Color <- ifelse(grepl("control", sample_RPM_GluGly_5p_tRF_long$Condition), "Control", paste0(sample_RPM_GluGly_5p_tRF_long$Group,"_Infected"))
```

```{r}
sample_RPM_GluGly_5p_tRF_long$Group <- factor(sample_RPM_GluGly_5p_tRF_long$Group, levels = c("T1_acute", "T2_acute", "T1_mild", "T2_mild"))
state.labs <- c("Severe (T1)", "Severe (T2)", "Moderate (T1)", "Moderate (T2)")
names(state.labs) <- c("T1_acute", "T2_acute", "T1_mild", "T2_mild")
sample_RPM_GluGly_5p_tRF_long$Label <- gsub("T[012]\\_(acute|mild|control)-", "", sample_RPM_GluGly_5p_tRF_long$Sample)

ggplot(sample_RPM_GluGly_5p_tRF_long, aes(x=Condition, y=RPM, fill=Color, color=Color)) +
  geom_violin(alpha=0.7) +
  geom_boxplot(width=0.25, outlier.shape = NA, alpha=0.7, color="black") +
  geom_jitter(position=position_jitter(0.2), alpha=0.7) +
  scale_fill_manual(values=c("#00468BB2", "#da6a65", "lightgreen", "#ED0000B2", "darkgreen")) +
  scale_color_manual(values=c("#00468BB2", "#da6a65", "lightgreen","#ED0000B2", "darkgreen")) +
  #geom_text_repel() +
  theme_classic() +
  facet_wrap(~Group, nrow = 1, strip.position = "bottom", labeller = labeller(Group=state.labs)) +
  theme(legend.position="none",
        axis.title = element_text(size=15),
        plot.title = element_text(size = rel(2)),
        axis.text.x = element_text(angle=45, vjust=1, hjust=1, color="black", size=12),
        axis.text.y = element_text(color="black", size=12),
        panel.grid = element_blank(),
        strip.background = element_rect(fill="#ebeff1"),
        strip.text = element_text(size=12),
        panel.border = element_blank(),
        axis.line.y = element_line(color="black"),
        axis.line.x = element_blank()) + ggtitle("5' tRF") + xlab("") + ylim(c(-1,6000))
ggsave("/path/to/tRNA_results/GluGly_violin_5p_tRF_withlabel.png", height = 5, width = 7)
```

#### 3-tRF

```{r}
sample_RPM_GluGly_3p_tRF <- data.frame(Sample=sampletable$SampleName[-c(1:8)], Group=sampletable$Group[-c(1:8)], Glu_control=NA, 
                                          Glu_infected=NA, Gly_control=NA, Gly_infected=NA)

for (case in c("T1_acute", "T1_mild", "T2_acute", "T2_mild")){
  for (aa in c("Glu", "Gly")){
    df <- get(paste0("DE_", case, "_tRNA"))
    df_subset <- df[df$ID_DB_cut == paste0("tRNA-",aa) & df$tRNA_class == "3p-tRF",]
    seqs <- df_subset$seq
    col_control <- colnames(RPMs_table)[which(grepl("control", colnames(RPMs_table)))]
    col_case <- colnames(RPMs_table)[which(grepl(case, colnames(RPMs_table)))]
    df_RPM <- RPMs_table[RPMs_table$seq %in% seqs, c("seq",col_control,col_case)]
    df_RPM <- df_RPM %>% #Sum every seq
      bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"Total")))
    sample_RPM_GluGly_3p_tRF[sample_RPM_GluGly_3p_tRF$Group == case, paste0(aa, "_control")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl("control", colnames(df_RPM))])
    sample_RPM_GluGly_3p_tRF[sample_RPM_GluGly_3p_tRF$Group == case, paste0(aa, "_infected")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl(case, colnames(df_RPM))])
  }
}

sample_RPM_GluGly_3p_tRF_long <- sample_RPM_GluGly_3p_tRF%>%
  pivot_longer(c(3:6),
               names_to="Condition",
               values_to="RPM")
sample_RPM_GluGly_3p_tRF_long$Color <- ifelse(grepl("control", sample_RPM_GluGly_3p_tRF_long$Condition), "Control", paste0(sample_RPM_GluGly_3p_tRF_long$Group,"_Infected"))
```

```{r}
sample_RPM_GluGly_3p_tRF_long$Group <- factor(sample_RPM_GluGly_3p_tRF_long$Group, levels = c("T1_acute", "T2_acute", "T1_mild", "T2_mild"))
state.labs <- c("Severe (T1)", "Severe (T2)", "Moderate (T1)", "Moderate (T2)")
names(state.labs) <- c("T1_acute", "T2_acute", "T1_mild", "T2_mild")
sample_RPM_GluGly_3p_tRF_long$Label <- gsub("T[012]\\_(acute|mild|control)-", "", sample_RPM_GluGly_3p_tRF_long$Sample)

ggplot(sample_RPM_GluGly_3p_tRF_long, aes(x=Condition, y=RPM, fill=Color, color=Color)) +
  geom_violin(alpha=0.7) +
  geom_boxplot(width=0.25, outlier.shape = NA, alpha=0.7, color="black") +
  geom_jitter(position=position_jitter(0.2), alpha=0.7) +
  scale_fill_manual(values=c("#00468BB2", "#da6a65", "lightgreen", "#ED0000B2", "darkgreen")) +
  scale_color_manual(values=c("#00468BB2", "#da6a65", "lightgreen","#ED0000B2", "darkgreen")) +
  #geom_text_repel() +
  theme_classic() +
  facet_wrap(~Group, nrow = 1, strip.position = "bottom", labeller = labeller(Group=state.labs)) +
  theme(legend.position="none",
        axis.title = element_text(size=15),
        plot.title = element_text(size = rel(2)),
        axis.text.x = element_text(angle=45, vjust=1, hjust=1, color="black", size=12),
        axis.text.y = element_text(color="black", size=12),
        panel.grid = element_blank(),
        strip.background = element_rect(fill="#ebeff1"),
        strip.text = element_text(size=12),
        panel.border = element_blank(),
        axis.line.y = element_line(color="black"),
        axis.line.x = element_blank()) + ggtitle("3' tRF") + xlab("") + ylim(c(-1,100))
ggsave("/path/to/tRNA_results/GluGly_violin_3p_tRF_withlabels.png", height = 5, width = 7)
```


***

### Supp. figure

We need an initial figure showing that the preponderant tRNA species (in terms of accumulation) are tRNA-Glu and tRNA-Gly.

####Severe
```{r}
sample_RPM_acute <- data.frame(Sample=sampletable$SampleName[c(9:16, 25:32)], Group=sampletable$Group[c(9:16, 25:32)])
extra_colnames <- paste0(rep(aa_list, each=2), c("_control", "_infected"))
df_extra = data.frame(matrix(nrow = nrow(sample_RPM_acute), ncol = length(extra_colnames))) 
colnames(df_extra) = extra_colnames
sample_RPM_acute <- cbind(sample_RPM_acute, df_extra)

for (case in c("T1_acute", "T2_acute")){
  for (aa in aa_list) {
    df <- get(paste0("DE_", case, "_tRNA"))
    df_subset <- df[df$ID_DB_cut == paste0("tRNA-",aa),]
    seqs <- df_subset$seq
    col_control <- colnames(RPMs_table)[which(grepl("control", colnames(RPMs_table)))]
    col_case <- colnames(RPMs_table)[which(grepl(case, colnames(RPMs_table)))]
    df_RPM <- RPMs_table[RPMs_table$seq %in% seqs, c("seq",col_control,col_case)]
    df_RPM <- df_RPM %>% #Sum every seq
      bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"Total")))
    sample_RPM_acute[sample_RPM_acute$Group == case, paste0(aa, "_control")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl("control", colnames(df_RPM))])
    sample_RPM_acute[sample_RPM_acute$Group == case, paste0(aa, "_infected")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl(case, colnames(df_RPM))])
  }
}

sample_RPM_acute_long <- sample_RPM_acute%>%
  pivot_longer(c(3:38),
               names_to="Condition",
               values_to="RPM")
sample_RPM_acute_long$tRNA <- str_replace(sample_RPM_acute_long$Condition, "_(control|infected)", "")
sample_RPM_acute_long$Group_toplot <- NA
sample_RPM_acute_long$Group_toplot <- ifelse(grepl("T1", sample_RPM_acute_long$Group) & grepl("control", sample_RPM_acute_long$Condition), 
                                             "Control_T1", sample_RPM_acute_long$Group_toplot)
sample_RPM_acute_long$Group_toplot <- ifelse(grepl("T1", sample_RPM_acute_long$Group) & grepl("infected", sample_RPM_acute_long$Condition), 
                                             "Infected_T1", sample_RPM_acute_long$Group_toplot)
sample_RPM_acute_long$Group_toplot <- ifelse(grepl("T2", sample_RPM_acute_long$Group) & grepl("control", sample_RPM_acute_long$Condition), 
                                             "Control_T2", sample_RPM_acute_long$Group_toplot)
sample_RPM_acute_long$Group_toplot <- ifelse(grepl("T2", sample_RPM_acute_long$Group) & grepl("infected", sample_RPM_acute_long$Condition), 
                                             "Infected_T2", sample_RPM_acute_long$Group_toplot)
#sample_RPM_acute_long$Color <- ifelse(grepl("control", sample_RPM_GluGly_3p_tRF_long$Condition), "Control", paste0(sample_RPM_GluGly_3p_tRF_long$Group,"_Infected"))
```


```{r}
sample_RPM_acute_long$Group_toplot <- factor(sample_RPM_acute_long$Group_toplot, levels = c("Control_T1", "Infected_T1", "Control_T2", "Infected_T2"))
ggplot(sample_RPM_acute_long, aes(x=Group_toplot, y=RPM, fill=Group_toplot, color=Group_toplot)) +
  geom_boxplot(width=0.25, outlier.shape = NA, alpha=0.7, color="black") +
  geom_jitter(position=position_jitter(0.2), alpha=0.7) +
  scale_fill_manual(values=c("#00468BB2", "#da6a65", "#00468BB2", "#ED0000B2")) +
  scale_color_manual(values=c("#00468BB2", "#da6a65", "#00468BB2", "#ED0000B2")) +
  facet_wrap(~ tRNA, nrow = 1, strip.position = "bottom") +
  theme_classic() +
  theme(legend.position="none",
        axis.title = element_text(size = 14),
        plot.title = element_text(size = rel(2)),
        axis.text.x = element_text(angle=45, vjust=1, hjust=1, color="black", size=12),
        axis.text.y = element_text(color="black", size=12),
        panel.grid = element_blank(),
        strip.background = element_rect(fill="#ebeff1"),
        strip.text = element_text(size=12),
        panel.border = element_blank(),
        axis.line.y = element_line(color="black"),
        axis.line.x = element_blank()) + xlab("")

ggsave("/path/to/tRNA_results/tRNA_isotype_accumulation_severe.png", height=7, width=15)
```

The exact number of the medians facilitated in the paper:

```{r}
median(pull(sample_RPM_acute_long[sample_RPM_acute_long$tRNA == "Glu" & sample_RPM_acute_long$Group_toplot == "Infected_T1", "RPM"]))
median(pull(sample_RPM_acute_long[sample_RPM_acute_long$tRNA == "Glu" & sample_RPM_acute_long$Group_toplot == "Infected_T2", "RPM"]))
median(pull(sample_RPM_acute_long[sample_RPM_acute_long$tRNA == "Gly" & sample_RPM_acute_long$Group_toplot == "Infected_T1", "RPM"]))
median(pull(sample_RPM_acute_long[sample_RPM_acute_long$tRNA == "Gly" & sample_RPM_acute_long$Group_toplot == "Infected_T2", "RPM"]))
```

#### Moderate

```{r}
sample_RPM_mild <- data.frame(Sample=sampletable$SampleName[c(17:24, 33:40)], Group=sampletable$Group[c(17:24, 33:40)])
extra_colnames <- paste0(rep(aa_list, each=2), c("_control", "_infected"))
df_extra = data.frame(matrix(nrow = nrow(sample_RPM_mild), ncol = length(extra_colnames))) 
colnames(df_extra) = extra_colnames
sample_RPM_mild <- cbind(sample_RPM_mild, df_extra)

for (case in c("T1_mild", "T2_mild")){
  for (aa in aa_list) {
    df <- get(paste0("DE_", case, "_tRNA"))
    df_subset <- df[df$ID_DB_cut == paste0("tRNA-",aa),]
    seqs <- df_subset$seq
    col_control <- colnames(RPMs_table)[which(grepl("control", colnames(RPMs_table)))]
    col_case <- colnames(RPMs_table)[which(grepl(case, colnames(RPMs_table)))]
    df_RPM <- RPMs_table[RPMs_table$seq %in% seqs, c("seq",col_control,col_case)]
    df_RPM <- df_RPM %>% #Sum every seq
      bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"Total")))
    sample_RPM_mild[sample_RPM_mild$Group == case, paste0(aa, "_control")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl("control", colnames(df_RPM))])
    sample_RPM_mild[sample_RPM_mild$Group == case, paste0(aa, "_infected")] <- as.numeric(df_RPM[df_RPM$seq == "Total", grepl(case, colnames(df_RPM))])
  }
}

sample_RPM_mild_long <- sample_RPM_mild%>%
  pivot_longer(c(3:38),
               names_to="Condition",
               values_to="RPM")
sample_RPM_mild_long$tRNA <- str_replace(sample_RPM_mild_long$Condition, "_(control|infected)", "")
sample_RPM_mild_long$Group_toplot <- NA
sample_RPM_mild_long$Group_toplot <- ifelse(grepl("T1", sample_RPM_mild_long$Group) & grepl("control", sample_RPM_mild_long$Condition), 
                                             "Control_T1", sample_RPM_mild_long$Group_toplot)
sample_RPM_mild_long$Group_toplot <- ifelse(grepl("T1", sample_RPM_mild_long$Group) & grepl("infected", sample_RPM_mild_long$Condition), 
                                             "Infected_T1", sample_RPM_mild_long$Group_toplot)
sample_RPM_mild_long$Group_toplot <- ifelse(grepl("T2", sample_RPM_mild_long$Group) & grepl("control", sample_RPM_mild_long$Condition), 
                                             "Control_T2", sample_RPM_mild_long$Group_toplot)
sample_RPM_mild_long$Group_toplot <- ifelse(grepl("T2", sample_RPM_mild_long$Group) & grepl("infected", sample_RPM_mild_long$Condition), 
                                             "Infected_T2", sample_RPM_mild_long$Group_toplot)
#sample_RPM_mild_long$Color <- ifelse(grepl("control", sample_RPM_GluGly_3p_tRF_long$Condition), "Control", paste0(sample_RPM_GluGly_3p_tRF_long$Group,"_Infected"))
```


```{r}
sample_RPM_mild_long$Group_toplot <- factor(sample_RPM_mild_long$Group_toplot, levels = c("Control_T1", "Infected_T1", "Control_T2", "Infected_T2"))
ggplot(sample_RPM_mild_long, aes(x=Group_toplot, y=RPM, fill=Group_toplot, color=Group_toplot)) +
  geom_boxplot(width=0.25, outlier.shape = NA, alpha=0.7, color="black") +
  geom_jitter(position=position_jitter(0.2), alpha=0.7) +
  scale_fill_manual(values=c("#00468BB2", "lightgreen", "#00468BB2", "darkgreen")) +
  scale_color_manual(values=c("#00468BB2", "lightgreen", "#00468BB2","darkgreen")) +
  scale_y_break(c(125000,210000)) + 
  facet_wrap(~ tRNA, nrow = 1, strip.position = "bottom") +
  theme_classic() +
  theme(legend.position="none",
        axis.title = element_text(size = 14),
        plot.title = element_text(size = rel(2)),
        axis.text.x = element_text(angle=45, vjust=1, hjust=1, color="black", size=12),
        axis.text.y = element_text(color="black", size=12),
        panel.grid = element_blank(),
        strip.background = element_rect(fill="#ebeff1"),
        strip.text = element_text(size=12),
        panel.border = element_blank(),
        axis.line.y = element_line(color="black"),
        axis.line.x = element_blank()) + xlab("")

ggsave("/path/to/tRNA_results/tRNA_isotype_accumulation_moderate.png", height=7, width=15)
```

The exact number of the medians facilitated in the paper:

```{r}
median(pull(sample_RPM_mild_long[sample_RPM_mild_long$tRNA == "Glu" & sample_RPM_mild_long$Group_toplot == "Infected_T1", "RPM"]))
median(pull(sample_RPM_mild_long[sample_RPM_mild_long$tRNA == "Glu" & sample_RPM_mild_long$Group_toplot == "Infected_T2", "RPM"]))
median(pull(sample_RPM_mild_long[sample_RPM_mild_long$tRNA == "Gly" & sample_RPM_mild_long$Group_toplot == "Infected_T1", "RPM"]))
median(pull(sample_RPM_mild_long[sample_RPM_mild_long$tRNA == "Gly" & sample_RPM_mild_long$Group_toplot == "Infected_T2", "RPM"]))
```




