---
title: "PCA_v3"
author: "Julia Corell Sierra"
date: "2024-02-20"
output: html_document
---

The aim of this script is to build a PCA containing all the samples of the SARS-CoV-2 study, i.e. for the different groups (control, acute and mild) and for the different times (T0, T1 and T2) all replicates must be present. As input data, we will use the same table that is used for DEA: the matrix of absolute counts (filtered considering only the sequences with 5 counts in at least five samples of any of the groups) -> Filtered_abs-outer_filt5.tsv

### 0. LOAD THE LIBRARIES AND SET THE OPTIONS

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
#options(rgl.useNULL=TRUE)
library(rgl)
#options(rgl.printRglwidget = TRUE)
setupKnitr(autoprint = TRUE)
knit_hooks$set(webgl = hook_webgl)
knitr.duplicate.label = "allow"
```

```{r}
library(tidyverse)
library(stats)
library(factoextra)
library(ggfortify)
library(DESeq2)
library(btools)
library(RColorBrewer)
library(pheatmap)
```

```{r}
sampletable <- read.table("../../Additional_data/samplesheet.tsv", header=T, sep="\t")

counts_table <- read.table("/path/to/results/03-Fusion_tables/Filtered_abs-outer_filt5.tsv",  header=T, sep="\t")
counts_table <- counts_table %>%
  column_to_rownames(var="seq")

```

Following the vignette of DESeq2 "Many common statistical methods for exploratory analysis of multidimensional data, for example clustering and principal components analysis (PCA), work best for data that generally has the same range of variance at different ranges of the mean values. [...] As a solution, DESeq2 offers two transformations for count data that stabilize the variance across the mean: the variance stabilizing transformation (VST) for negative binomial data with a dispersion-mean trend (Anders and Huber 2010), implemented in the vst function, and the regularized-logarithm transformation or rlog (Love, Huber, and Anders 2014)."
```{r}
sampletable$Age_group <- NA
sampletable$Age_group <- cut(sampletable$Age, 4, labels = FALSE)

dds <- DESeqDataSetFromMatrix(countData = counts_table,
                                 colData = sampletable,
                                 design = ~ Group)
#rm(counts_table)
dds$Age_group <- factor(dds$Age_group, c(1,2,3,4))
dds$Group <- relevel(dds$Group, ref ="T0_control")
#dds2 <- DESeq(dds)
vst_dds <- assay(vst(dds, blind=TRUE))
vst_3d <- vst(dds, blind=TRUE)

# As we are now working with the outer table, one way to filter the rows is
# to work only with those N sequences with highest row variance. Such as they
# do in DESeq2::plotPCA.
rv <- rowVars(vst_dds)
select <- order(rv, decreasing = TRUE)[seq_len(min(1000,
            length(rv)))]
pca_res <- prcomp(x = t(vst_dds[select, ]), rank. = 6)
pca_df <- as.data.frame(pca_res$x)
```


```{r}
fviz_eig(pca_res, addlabels=TRUE)
```


```{r}
## Convert to dataframe and add a color column and a names columns (sample names)
df_vst = as.data.frame(pca_df)

## Add the color.
df_vst$color = c(rep("#00468BB2",8), #T0_control --> Azul.
              rep("#da6a65",8), #T1_acute --> Rojo claro
              rep("lightgreen",8), #T1_mild --> Verde claro
              rep("#ED0000B2",8), #T2_acute --> Rojo oscuro
              rep("darkgreen",8)) #T2_mild --> Verde oscuro
        

## Plot the PCA
with(df_vst, par3d(ignoreExtent=F))
with(df_vst,plot3d(df_vst$PC1, df_vst$PC2, df_vst$PC3, col=df_vst$color,type="s", radius=6, alpha=1, 
               xlab = "", ylab = "", zlab = "", shininess=20, axes=F))

with(df_vst, box3d())
with(df_vst, axis3d('x--',labels=T,tick=T))
with(df_vst, axis3d('y+-',labels=T,tick=T))
with(df_vst, axis3d('z++',labels=T,tick=T))

with(df_vst, par3d(ignoreExtent=T))
with(df_vst,aspect3d(1, 1, 1))
with(df_vst,title3d(main = '', sub = NULL, 
                xlab = "PC1", ylab = "PC2", color = "black",
                zlab = "PC3", font = 2))

```

We will perform additional sample-to-sample distance visualization to study the effect of variables such as sex, sampling intervals or age. 

```{r}
sampleDists <- dist(t(assay(vst(dds, blind=TRUE))))
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- paste(vst_3d$SampleName)
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
row_anno <- data.frame(
  rownames = sampletable$SampleName,
  Sex = sampletable$Sex,
  Age = sampletable$Age_group,
  Condition = sampletable$Condition
)
rownames(row_anno) <- row_anno[,1]
row_anno <- row_anno[-1]
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors,
         annotation_row = row_anno)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors,
         annotation_row = row_anno)
```


```{r}
counts_table_nocont <- counts_table[, !grepl("control", colnames(counts_table))]
sampletable_nocont <- sampletable %>%
  filter(Condition != "Control")
dds_nocont <- DESeqDataSetFromMatrix(countData = counts_table_nocont,
                                 colData = sampletable_nocont,
                                 design = ~ Group)
dds_nocont$Age_group <- factor(dds_nocont$Age_group, c(1,2,3,4))
vst_3d_nocont <- vst(dds_nocont, blind=TRUE)
vst_dds_nocont <- assay(vst(dds_nocont, blind=TRUE)) # To extract PC %

# As we are now working with the outer table, one way to filter the rows is
# to work only with those N sequences with highest row variance. Such as they
# do in DESeq2::plotPCA.
rv_nocont <- rowVars(vst_dds_nocont)
select_nocont <- order(rv_nocont, decreasing = TRUE)[seq_len(min(1000,
            length(rv_nocont)))]
```
```{r}
sampleDists_nocont <- dist(t(assay(vst_3d_nocont[select_nocont,])))
sampleDistMatrix_nocont <- as.matrix(sampleDists_nocont)
rownames(sampleDistMatrix_nocont) <- paste(vst_3d_nocont$SampleName)
colnames(sampleDistMatrix_nocont) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
row_anno <- data.frame(
  rownames = sampletable_nocont$SampleName,
  Sex = sampletable_nocont$Sex,
  Age = sampletable_nocont$Age_group,
  Condition = sampletable_nocont$Condition
)
rownames(row_anno) <- row_anno[,1]
row_anno <- row_anno[-1]
pheatmap(sampleDistMatrix_nocont,
         clustering_distance_rows=sampleDists_nocont,
         clustering_distance_cols=sampleDists_nocont,
         col=colors,
         annotation_row = row_anno,
         filename = "/path/to/results/05-PCA/sampleDist_T1andT2.png")
```

It may be a good idea to plot T1 and T2 separatedly.

```{r}
counts_table_T1 <- counts_table[, grepl("T1", colnames(counts_table))]
sampletable_T1 <- sampletable %>%
  filter(Time == "T1")

dds_T1 <- DESeqDataSetFromMatrix(countData = counts_table_T1,
                                 colData = sampletable_T1,
                                 design = ~ Group)

vst_3d_T1 <- vst(dds_T1, blind=TRUE)
vst_dds_T1 <- assay(vst(dds_T1, blind=TRUE)) # To extract PC %

rv_T1 <- rowVars(vst_dds_T1)
select_T1 <- order(rv_T1, decreasing = TRUE)[seq_len(min(1000,
            length(rv_T1)))]

sampleDists_T1 <- dist(t(assay(vst_3d_T1[select_T1,])))
sampleDistMatrix_T1 <- as.matrix(sampleDists_T1)
rownames(sampleDistMatrix_T1) <- paste(vst_3d_T1$SampleName)
colnames(sampleDistMatrix_T1) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
row_anno_T1 <- data.frame(
  rownames = sampletable_T1$SampleName,
  Sex = sampletable_T1$Sex,
  Age = sampletable_T1$Age_group,
  Condition = sampletable_T1$Condition
)
rownames(row_anno_T1) <- row_anno_T1[,1]
row_anno_T1 <- row_anno_T1[-1]
pheatmap(sampleDistMatrix_T1,
         clustering_distance_rows=sampleDists_T1,
         clustering_distance_cols=sampleDists_T1,
         col=colors,
         annotation_row = row_anno_T1,
         filename = "/path/to/results/05-PCA/sampleDist_T1.png",
         width = 8,
         height = 6)
```



Let's explore if the time interval might have some influence. Plot only T2 samples.

```{r}
counts_table_T2 <- counts_table[, grepl("T2", colnames(counts_table))]
sampletable_T2 <- sampletable %>%
  filter(Time == "T2")
sampletable_T2$Interval_group <- cut(sampletable_T2$Interval, 4, labels = FALSE)
dds_T2 <- DESeqDataSetFromMatrix(countData = counts_table_T2,
                                 colData = sampletable_T2,
                                 design = ~ Group)
dds_T2$Interval_group <- factor(dds_T2$Interval_group, c(1,2,3,4))
vst_3d_T2 <- vst(dds_T2, blind=TRUE)
vst_dds_T2 <- assay(vst(dds_T2, blind=TRUE)) # To extract PC %

# As we are now working with the outer table, one way to filter the rows is
# to work only with those N sequences with highest row variance. Such as they
# do in DESeq2::plotPCA.
rv_T2 <- rowVars(vst_dds_T2)
select_T2 <- order(rv_T2, decreasing = TRUE)[seq_len(min(1000,
            length(rv_T2)))]

sampleDists_T2 <- dist(t(assay(vst_3d_T2[select_T2,])))
sampleDistMatrix_T2 <- as.matrix(sampleDists_T2)
rownames(sampleDistMatrix_T2) <- paste(vst_3d_T2$SampleName)
colnames(sampleDistMatrix_T2) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
row_anno_T2 <- data.frame(
  rownames = sampletable_T2$SampleName,
  Sex = sampletable_T2$Sex,
  Age = sampletable_T2$Age_group,
  Sampling_interval = sampletable_T2$Interval_group,
  Condition = sampletable_T2$Condition
)
rownames(row_anno_T2) <- row_anno_T2[,1]
row_anno_T2 <- row_anno_T2[-1]
pheatmap(sampleDistMatrix_T2,
         clustering_distance_rows=sampleDists_T2,
         clustering_distance_cols=sampleDists_T2,
         col=colors,
         annotation_row = row_anno_T2,
         filename = "/path/to/results/05-PCA/sampleDist_T2.png",
         width = 8,
         height = 6)
```
