---
title: "miRNA and tyRNA analysis"
author: "Julia Corell Sierra"
date: "2023-06-12"
output: html_document
---


In this script you'll find all the code necessary to produce the figures and tables related to the miRNA and tyRNA analysis.
```{r}
library("Biostrings")
library("ggplot2")
library("tidyverse")
library("ggforestplot")
```

```{r}
padj.cutoff <- 0.05
anno_file <- read.table("/path/to/annotation_results/unitas_simplified_annnotation/curated_unitas_withtiny_annotation.tsv", sep="\t", header=TRUE)
DE_T1_acute <- read.table("/path/to/DEA_results/results_DE_T1acute.tsv", sep="\t", header=TRUE)
DE_T1_acute_sig <- DE_T1_acute %>%
  dplyr::filter(padj < padj.cutoff)
DE_T1_mild <- read.table("/path/to/DEA_results/results_DE_T1mild.tsv", sep="\t", header=TRUE)
DE_T1_mild_sig <- DE_T1_mild %>%
  dplyr::filter(padj < padj.cutoff)
DE_T2_acute <- read.table("/path/to/DEA_results/results_DE_T2acute.tsv", sep="\t", header=TRUE)
DE_T2_acute_sig <- DE_T2_acute %>%
  dplyr::filter(padj < padj.cutoff)
DE_T2_mild <- read.table("/path/to/DEA_results/results_DE_T2mild.tsv", sep="\t", header=TRUE)
DE_T2_mild_sig <- DE_T2_mild %>%
  dplyr::filter(padj < padj.cutoff)
DE_T1_acute_sig_ann <- merge(DE_T1_acute_sig, anno_file, by.x = "seq", by.y = "Seq", all.x = TRUE)
DE_T1_mild_sig_ann <-  merge(DE_T1_mild_sig, anno_file, by.x = "seq", by.y = "Seq", all.x = TRUE)
DE_T2_acute_sig_ann <- merge(DE_T2_acute_sig, anno_file, by.x = "seq", by.y = "Seq", all.x = TRUE)
DE_T2_mild_sig_ann <-  merge(DE_T2_mild_sig, anno_file, by.x = "seq", by.y = "Seq", all.x = TRUE)
```


```{r}
DE_T1_acute_miR <- DE_T1_acute_sig_ann %>%
  filter(grepl("miRNA$", RNA_class))
DE_T1_acute_miR$Family <- DE_T1_acute_miR$RNA_ID

DE_T1_acute_miR$Family <- str_remove(DE_T1_acute_miR$Family, "-[35]p")
DE_T1_acute_miR$Family <- str_remove(DE_T1_acute_miR$Family, "\\(.+\\)")
DE_T1_acute_miR$Family <- str_remove(DE_T1_acute_miR$Family, "\\-[1-9]$")
DE_T1_acute_miR$Family <- str_remove(DE_T1_acute_miR$Family, "[a-z]{1,}$")
DE_T1_acute_miR$Family <- paste0("hsa-", DE_T1_acute_miR$Family)
```


```{r}
DE_T1_mild_miR <- DE_T1_mild_sig_ann %>%
  filter(grepl("miRNA$", RNA_class))
DE_T1_mild_miR$Family <- DE_T1_mild_miR$RNA_ID

DE_T1_mild_miR$Family <- str_remove(DE_T1_mild_miR$Family, "-[35]p")
DE_T1_mild_miR$Family <- str_remove(DE_T1_mild_miR$Family, "\\(.+\\)")
DE_T1_mild_miR$Family <- str_remove(DE_T1_mild_miR$Family, "\\-[1-9]$")
DE_T1_mild_miR$Family <- str_remove(DE_T1_mild_miR$Family, "[a-z]{1,}$")
DE_T1_mild_miR$Family <- paste0("hsa-", DE_T1_mild_miR$Family)
```


```{r}
DE_T2_acute_miR <- DE_T2_acute_sig_ann %>%
  filter(grepl("miRNA$", RNA_class))
DE_T2_acute_miR$Family <- DE_T2_acute_miR$RNA_ID

DE_T2_acute_miR$Family <- str_remove(DE_T2_acute_miR$Family, "-[35]p")
DE_T2_acute_miR$Family <- str_remove(DE_T2_acute_miR$Family, "\\(.+\\)")
DE_T2_acute_miR$Family <- str_remove(DE_T2_acute_miR$Family, "\\-[1-9]$")
DE_T2_acute_miR$Family <- str_remove(DE_T2_acute_miR$Family, "[a-z]{1,}$")
DE_T2_acute_miR$Family <- paste0("hsa-", DE_T2_acute_miR$Family)
```


```{r}
DE_T2_mild_miR <- DE_T2_mild_sig_ann %>%
  filter(grepl("miRNA$", RNA_class))
DE_T2_mild_miR$Family <- DE_T2_mild_miR$RNA_ID

DE_T2_mild_miR$Family <- str_remove(DE_T2_mild_miR$Family, "-[35]p")
DE_T2_mild_miR$Family <- str_remove(DE_T2_mild_miR$Family, "\\(.+\\)")
DE_T2_mild_miR$Family <- str_remove(DE_T2_mild_miR$Family, "\\-[1-9]$")
DE_T2_mild_miR$Family <- str_remove(DE_T2_mild_miR$Family, "[a-z]{1,}$")
DE_T2_mild_miR$Family <- paste0("hsa-", DE_T2_mild_miR$Family)
```


***
#### miRNA analysis 

Supplemental figure:

Plot of only those miRNAs annotated with high confidence (without mismatch).

```{r}
DE_T1_acute_miR_highconf <- DE_T1_acute_miR[!grepl("\\(", DE_T1_acute_miR$RNA_ID),]
DE_T2_acute_miR_highconf <- DE_T2_acute_miR[!grepl("\\(", DE_T2_acute_miR$RNA_ID),]
DE_T1_mild_miR_highconf <- DE_T1_mild_miR[!grepl("\\(", DE_T1_mild_miR$RNA_ID),]
DE_T2_mild_miR_highconf <- DE_T2_mild_miR[!grepl("\\(", DE_T2_mild_miR$RNA_ID),]
```

Make the plots for each time and disease severity comparable by showing all miRNAs DE.

```{r}
families_highconf <- Reduce(union, list(DE_T1_acute_miR_highconf$Family, DE_T1_mild_miR_highconf$Family, DE_T2_acute_miR_highconf$Family, DE_T2_mild_miR_highconf$Family))
families_highconf <- sort(families_highconf)
```

```{r}
DE_T1_acute_miR_highconf$Family <- factor(DE_T1_acute_miR_highconf$Family, levels = families_highconf)
ggplot(DE_T1_acute_miR_highconf, aes(x=Family, y=lfcShrunk)) +
  geom_boxplot(outlier.shape=NA, fill="#da6a65", alpha=0.7) +
  geom_jitter(size=0.8, shape=16, alpha=0.7) +
  theme_bw() +
  scale_x_discrete(drop=FALSE) + 
  geom_hline(yintercept=0, color="black") +
  theme(legend.position = "none",
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(color="black", angle=90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(color="black")) + ggtitle("Severe (T1) miRNA family expression") + ylab("Log2 fold change") +
  xlab("miRNA family")

ggsave("/path/to/miRNA_and_tiny_results/T1_acute_miR_all_families_highconf.png", height=5, width=17, dpi=600)
```

```{r}
DE_T1_mild_miR_highconf$Family <- factor(DE_T1_mild_miR_highconf$Family, levels = families_highconf)
ggplot(DE_T1_mild_miR_highconf, aes(x=Family, y=lfcShrunk)) +
  geom_boxplot(outlier.shape=NA, fill="lightgreen", alpha=0.7) +
  geom_jitter(size=0.8, shape=16, alpha=0.7) +
  scale_x_discrete(drop=FALSE) +
  theme_bw() +
  geom_hline(yintercept=0, color="black") +
  theme(legend.position = "none",
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(color="black", angle=90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(color="black")) + ggtitle("Moderate (T1) miRNA family expression") + ylab("Log2 fold change") +
  xlab("miRNA family")
ggsave("/path/to/miRNA_and_tiny_results/T1_mild_miR_all_families_highconf.png", height=5, width=17, dpi=600)
```

```{r}
DE_T2_acute_miR_highconf$Family <- factor(DE_T2_acute_miR_highconf$Family, levels = families_highconf)
ggplot(DE_T2_acute_miR_highconf, aes(x=Family, y=lfcShrunk)) +
  geom_boxplot(outlier.shape=NA, fill="#ED0000B2", alpha=0.7) +
  geom_jitter(size=0.8, shape=16, alpha=0.7) +
  scale_x_discrete(drop=FALSE) +
  theme_bw() +
  geom_hline(yintercept=0, color="black") +
  theme(legend.position = "none",
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(color="black", angle=90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(color="black")) + ggtitle("Severe (T2) miRNA family expression") + ylab("Log2 fold change") +
  xlab("miRNA family")

ggsave("/path/to/miRNA_and_tiny_results/T2_acute_miR_all_families_highconf.png", height=5, width=17, dpi=600)
```
```{r}
DE_T2_mild_miR_highconf$Family <- factor(DE_T2_mild_miR_highconf$Family, levels = families_highconf)
ggplot(DE_T2_mild_miR_highconf, aes(x=Family, y=lfcShrunk)) +
  geom_boxplot(outlier.shape=NA, fill="darkgreen", alpha=0.7) +
  geom_jitter(size=0.8, shape=16, alpha=0.7) +
  scale_x_discrete(drop=FALSE) +
  theme_bw() +
  geom_hline(yintercept=0, color="black") +
  theme(legend.position = "none",
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(color="black", angle=90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(color="black")) + ggtitle("Moderate (T2) miRNA family expression") + ylab("Log2 fold change") +
  xlab("miRNA family")

ggsave("/path/to/miRNA_and_tiny_results/T2_mild_miR_all_families_highconf.png", height=5, width=17, dpi=600)
```


#### Tables

First let's get a binary table of presence/absence of miRNA family per contrast.

```{r}
miRNA_binary_highconf <- data.frame(Family=families_highconf, T1_acute=NA, T2_acute=NA, T1_mild=NA, T2_mild=NA)
miRNA_binary_highconf$T1_acute <- ifelse(miRNA_binary_highconf$Family %in% unique(DE_T1_acute_miR_highconf$Family), 1, 0)
miRNA_binary_highconf$T2_acute <- ifelse(miRNA_binary_highconf$Family %in% unique(DE_T2_acute_miR_highconf$Family), 1, 0)
miRNA_binary_highconf$T1_mild <- ifelse(miRNA_binary_highconf$Family %in% unique(DE_T1_mild_miR_highconf$Family), 1, 0)
miRNA_binary_highconf$T2_mild <- ifelse(miRNA_binary_highconf$Family %in% unique(DE_T2_mild_miR_highconf$Family), 1, 0)

write.table(miRNA_binary_highconf, "/path/to/miRNA_and_tiny_results/miRNA_binary_table_highconf.tsv", sep="\t", row.names = FALSE)
```

Diversity of sequences per family and contrast.
```{r}
miRNA_seq_diversity_highconf <- data.frame(Family=families_highconf, T1_acute=NA, T2_acute=NA, T1_mild=NA, T2_mild=NA)

for (fam in families_highconf){
  miRNA_seq_diversity_highconf$T1_acute[miRNA_seq_diversity_highconf$Family == fam] <- nrow(DE_T1_acute_miR_highconf[DE_T1_acute_miR_highconf$Family == fam,])
  miRNA_seq_diversity_highconf$T2_acute[miRNA_seq_diversity_highconf$Family == fam] <- nrow(DE_T2_acute_miR_highconf[DE_T2_acute_miR_highconf$Family == fam,])
  miRNA_seq_diversity_highconf$T1_mild[miRNA_seq_diversity_highconf$Family == fam] <- nrow(DE_T1_mild_miR_highconf[DE_T1_mild_miR_highconf$Family == fam,])
  miRNA_seq_diversity_highconf$T2_mild[miRNA_seq_diversity_highconf$Family == fam] <- nrow(DE_T2_mild_miR_highconf[DE_T2_mild_miR_highconf$Family == fam,])
}

write.table(miRNA_seq_diversity_highconf, "/path/to/miRNA_and_tiny_results/miRNA_seqdiversity_table_highconf.tsv", sep="\t", row.names = FALSE)
```

Median LFC per family and contrast.
```{r}
miRNA_medianLFC_highconf <- data.frame(Family=families_highconf, T1_acute=NA, T2_acute=NA, T1_mild=NA, T2_mild=NA)

for (fam in families_highconf){
  miRNA_medianLFC_highconf$T1_acute[miRNA_medianLFC_highconf$Family == fam] <- ifelse(nrow(DE_T1_acute_miR_highconf[DE_T1_acute_miR_highconf$Family == fam,]) > 1 ,median(DE_T1_acute_miR_highconf[DE_T1_acute_miR_highconf$Family == fam, "lfcShrunk"]), DE_T1_acute_miR_highconf[DE_T1_acute_miR_highconf$Family == fam, "lfcShrunk"])
  miRNA_medianLFC_highconf$T2_acute[miRNA_medianLFC_highconf$Family == fam] <- ifelse(nrow(DE_T2_acute_miR_highconf[DE_T2_acute_miR_highconf$Family == fam,]) > 1 , median(DE_T2_acute_miR_highconf[DE_T2_acute_miR_highconf$Family == fam, "lfcShrunk"]), DE_T2_acute_miR_highconf[DE_T2_acute_miR_highconf$Family == fam, "lfcShrunk"])
  miRNA_medianLFC_highconf$T1_mild[miRNA_medianLFC_highconf$Family == fam] <- ifelse(nrow(DE_T1_mild_miR_highconf[DE_T1_mild_miR_highconf$Family == fam,]) > 1 , median(DE_T1_mild_miR_highconf[DE_T1_mild_miR_highconf$Family == fam, "lfcShrunk"]), DE_T1_mild_miR_highconf[DE_T1_mild_miR_highconf$Family == fam, "lfcShrunk"])
  miRNA_medianLFC_highconf$T2_mild[miRNA_medianLFC_highconf$Family == fam] <- ifelse(nrow(DE_T2_mild_miR_highconf[DE_T2_mild_miR_highconf$Family == fam,]) > 1 , median(DE_T2_mild_miR_highconf[DE_T2_mild_miR_highconf$Family == fam, "lfcShrunk"]), DE_T2_mild_miR_highconf[DE_T2_mild_miR_highconf$Family == fam, "lfcShrunk"])
}

write.table(miRNA_medianLFC_highconf, "/path/to/miRNA_and_tiny_results/miRNA_medianLFC_highconf.tsv", sep="\t", row.names = FALSE)
```

Detect those sequences that don't have the same trend.

```{r}
all(DE_T1_acute_miR_highconf[DE_T1_acute_miR_highconf$Family == fam, "lfcShrunk"] < 1)

for (fam in families_highconf){
  if(all(DE_T1_acute_miR_highconf[DE_T1_acute_miR_highconf$Family == fam, "lfcShrunk"] < 0) | all(DE_T1_acute_miR_highconf[DE_T1_acute_miR_highconf$Family == fam, "lfcShrunk"] > 0)){}
  else{print(paste("T1 acute: ", fam))}
  if(all(DE_T2_acute_miR_highconf[DE_T2_acute_miR_highconf$Family == fam, "lfcShrunk"] < 0) | all(DE_T2_acute_miR_highconf[DE_T2_acute_miR_highconf$Family == fam, "lfcShrunk"] > 0)){}
  else{print(paste("T2 acute: ", fam))}
  if(all(DE_T1_mild_miR_highconf[DE_T1_mild_miR_highconf$Family == fam, "lfcShrunk"] < 0) | all(DE_T1_mild_miR_highconf[DE_T1_mild_miR_highconf$Family == fam, "lfcShrunk"] > 0)){}
  else{print(paste("T1 mild: ",fam))}
  if(all(DE_T2_mild_miR_highconf[DE_T2_mild_miR_highconf$Family == fam, "lfcShrunk"] < 0) | all(DE_T2_mild_miR_highconf[DE_T2_mild_miR_highconf$Family == fam, "lfcShrunk"] > 0)){}
  else{print(paste("T2 mild: ",fam))}
}
```

**Questions:**

How many sequences are reactive to the infection state?

```{r}
##Exclusively in severe
table(miRNA_binary_highconf$T1_acute == 1 & miRNA_binary_highconf$T2_acute ==1 & miRNA_binary_highconf$T1_mild == 0 & miRNA_binary_highconf$T2_mild == 0)
##Exclusively in moderate
table(miRNA_binary_highconf$T1_acute == 0 & miRNA_binary_highconf$T2_acute ==0 & miRNA_binary_highconf$T1_mild == 1 & miRNA_binary_highconf$T2_mild == 1)
```

And related to the time point?
```{r}
##Exclusively in T1
table(miRNA_binary_highconf$T1_acute == 1 & miRNA_binary_highconf$T2_acute == 0 & miRNA_binary_highconf$T1_mild == 1 & miRNA_binary_highconf$T2_mild == 0)
##Exclusively in T2
table(miRNA_binary_highconf$T1_acute == 0 & miRNA_binary_highconf$T2_acute ==1 & miRNA_binary_highconf$T1_mild == 0 & miRNA_binary_highconf$T2_mild == 1)
```

***

The supplemental table that summarizes miRNA diversity and differential expression along the four contrasts.

Summary table of sequences and corresponding families:

```{r}
DE_T1_acute_sel <- DE_T1_acute_miR_highconf[, c("seq", "Family")]
DE_T2_acute_sel <- DE_T2_acute_miR_highconf[, c("seq", "Family")]
DE_T1_mild_sel <- DE_T1_mild_miR_highconf[, c("seq", "Family")]
DE_T2_mild_sel <- DE_T2_mild_miR_highconf[, c("seq", "Family")]

summ_miR_fam <- rbind(DE_T1_acute_sel, DE_T2_acute_sel, DE_T1_mild_sel, DE_T2_mild_sel)
summ_miR_fam <- unique(summ_miR_fam)
```

```{r}
seqs_DE_HC <- Reduce(union, list(DE_T1_acute_miR_highconf$seq, DE_T2_acute_miR_highconf$seq, DE_T1_mild_miR_highconf$seq, DE_T2_mild_miR_highconf$seq))

summary_miRNA_table <- data.frame(seq=seqs_DE_HC, miRNA_family=NA, Severe_T1_LFC=NA, Severe_T1_padj=NA, 
                                  Severe_T2_LFC=NA, Severe_T2_padj=NA, Moderate_T1_LFC=NA, Moderate_T1_padj=NA,
                                  Moderate_T2_LFC=NA, Moderate_T2_padj=NA)

#We need to unfactor the column to extract the values
summ_miR_fam$Family <- as.character(summ_miR_fam$Family)

for (sequence in seqs_DE_HC) {
  #Get miRNA family from the summary table
  summary_miRNA_table[summary_miRNA_table$seq == sequence, "miRNA_family"] <- summ_miR_fam[summ_miR_fam$seq == sequence, "Family"]
  summary_miRNA_table$Severe_T1_LFC[summary_miRNA_table$seq == sequence] <- ifelse(sequence %in% DE_T1_acute_miR_highconf$seq, DE_T1_acute_miR_highconf[DE_T1_acute_miR_highconf$seq == sequence, "lfcShrunk"], NA)
  summary_miRNA_table$Severe_T1_padj[summary_miRNA_table$seq == sequence] <- ifelse(sequence %in% DE_T1_acute_miR_highconf$seq, DE_T1_acute_miR_highconf[DE_T1_acute_miR_highconf$seq == sequence, "padj"], NA)
  summary_miRNA_table$Severe_T2_LFC[summary_miRNA_table$seq == sequence] <- ifelse(sequence %in% DE_T2_acute_miR_highconf$seq, DE_T2_acute_miR_highconf[DE_T2_acute_miR_highconf$seq == sequence, "lfcShrunk"], NA)
  summary_miRNA_table$Severe_T2_padj[summary_miRNA_table$seq == sequence] <- ifelse(sequence %in% DE_T2_acute_miR_highconf$seq, DE_T2_acute_miR_highconf[DE_T2_acute_miR_highconf$seq == sequence, "padj"], NA)
  summary_miRNA_table$Moderate_T1_LFC[summary_miRNA_table$seq == sequence] <- ifelse(sequence %in% DE_T1_mild_miR_highconf$seq, DE_T1_mild_miR_highconf[DE_T1_mild_miR_highconf$seq == sequence, "lfcShrunk"], NA) 
  summary_miRNA_table$Moderate_T1_padj[summary_miRNA_table$seq == sequence] <- ifelse(sequence %in% DE_T1_mild_miR_highconf$seq, DE_T1_mild_miR_highconf[DE_T1_mild_miR_highconf$seq == sequence, "padj"], NA)
  summary_miRNA_table$Moderate_T2_LFC[summary_miRNA_table$seq == sequence] <- ifelse(sequence %in% DE_T2_mild_miR_highconf$seq, DE_T2_mild_miR_highconf[DE_T2_mild_miR_highconf$seq == sequence, "lfcShrunk"], NA)
  summary_miRNA_table$Moderate_T2_padj[summary_miRNA_table$seq == sequence] <- ifelse(sequence %in% DE_T2_mild_miR_highconf$seq, DE_T2_mild_miR_highconf[DE_T2_mild_miR_highconf$seq == sequence, "padj"], NA)
}

summary_miRNA_table <- summary_miRNA_table[order(summary_miRNA_table$miRNA_family),]
write.table(summary_miRNA_table, "/path/to/miRNA_and_tiny_results/Supp_miRNA.tsv", sep="\t", row.names=FALSE)
```


***

For the principal miRNAs figure we need to get only the log2FC of those families that responded consistently in every group.

```{r}
miRNA_binary_highconf_filt <- miRNA_binary_highconf[apply(miRNA_binary_highconf[, -1], 1, function(x) all(x == 1)),]
miRNA_consist_fam <- miRNA_binary_highconf_filt$Family
#Remove hsa-miR-203
miRNA_consist_fam <- miRNA_consist_fam[-which(miRNA_consist_fam == "hsa-miR-203")]
```

```{r}
DE_T1_acute_toplot <- DE_T1_acute_miR_highconf[, c("seq", "lfcShrunk", "Family")]
DE_T1_acute_toplot <- DE_T1_acute_toplot %>% filter(Family %in% miRNA_consist_fam)
DE_T1_acute_toplot$Contrast <- "T1_acute"
DE_T1_acute_toplot$Severity <- "Severe"
DE_T1_acute_toplot$Time <- "T1"
```

```{r}
for (time in c("T1", "T2")){
  for (sev in c("acute", "mild")){
    df_cont <- get(paste0("DE_", time,"_", sev, "_miR_highconf"))
    df_cont_toplot <- df_cont[, c("seq", "lfcShrunk", "Family")]
    df_cont_toplot <- df_cont_toplot %>% filter(Family %in% miRNA_consist_fam)
    df_cont_toplot$Contrast <- paste0(time, "_", sev)
    df_cont_toplot$Severity <- sev
    df_cont_toplot$Time <- time
    name_df <- paste0("DE_", time, "_", sev, "_toplot")
    assign(name_df, df_cont_toplot)
  }
}
```

```{r}
miRNA_consist_toplot <- rbind(DE_T1_acute_toplot, DE_T2_acute_toplot, DE_T1_mild_toplot, DE_T2_mild_toplot)
miRNA_consist_toplot$Severity <- ifelse(miRNA_consist_toplot$Severity == "acute", "Severe", "Moderate")
miRNA_consist_toplot$Severity <- factor(miRNA_consist_toplot$Severity, levels=c("Severe", "Moderate"))
miRNA_consist_toplot$Contrast <- ifelse(miRNA_consist_toplot$Contrast == "T1_acute", "Severe (T1)", miRNA_consist_toplot$Contrast)
miRNA_consist_toplot$Contrast <- ifelse(miRNA_consist_toplot$Contrast == "T2_acute", "Severe (T2)", miRNA_consist_toplot$Contrast)
miRNA_consist_toplot$Contrast <- ifelse(miRNA_consist_toplot$Contrast == "T1_mild", "Moderate (T1)", miRNA_consist_toplot$Contrast)
miRNA_consist_toplot$Contrast <- ifelse(miRNA_consist_toplot$Contrast == "T2_mild", "Moderate (T2)", miRNA_consist_toplot$Contrast)
miRNA_consist_toplot$Contrast <- factor(miRNA_consist_toplot$Contrast, levels=c("Severe (T1)", "Severe (T2)", "Moderate (T1)", "Moderate (T2)"))
```


```{r}
ggplot(miRNA_consist_toplot, aes(x=Family, y=lfcShrunk, fill=Contrast)) +
  geom_boxplot(outlier.shape=NA, alpha=0.7) +
  geom_jitter(size=1.5, shape=16, alpha=0.7) +
  theme_bw() +
  geom_hline(yintercept=0, color="black") +
  theme(legend.position="bottom",
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(color="black"),
    axis.text.y = element_text(color="black")) + ylab("Log2 fold change") +
  facet_wrap(~ Severity, ncol=1) +
  scale_fill_manual(values=c("#da6a65", "#ED0000B2", "lightgreen",  "darkgreen")) + labs(fill="Group") + xlab("miRNA family")

ggsave("/path/to/miRNA_and_tiny_results/miR_consistent.png", height=5, width=8, dpi=600)
```


***

### TinyRNAs figure

```{r}
tiny_toplot <- read.table("/path/to/miRNA_and_tiny_results/tiny_table_toplot_filt.tsv", sep="\t", header=TRUE)
#tiny_toplot$Seq_name <- factor(tiny_toplot$Seq_name, levels=unique(tiny_toplot$Seq_name))
tiny_toplot[is.na(tiny_toplot)] <- 0
tiny_toplot$Color <- ifelse(tiny_toplot$LFC < 0, "blue", "red")
tiny_toplot$Color <- ifelse(tiny_toplot$LFC == 0, "grey", tiny_toplot$Color)
```

```{r}
rects <- data.frame(ystart = c(0,2.5,4.5,6.5,8.5,10.5,12.5,14.5), yend = c(2.5,4.5,6.5,8.5,10.5,12.5,14.5,16.5), col=(c("grey", "white", "grey", "white", "grey", "white", "grey", "white")))
```

```{r}
tiny_toplot_acute <- tiny_toplot[tiny_toplot$Group == "Severe",]
tiny_toplot_acute$Seq_name <- factor(tiny_toplot_acute$Seq_name, levels=unique(tiny_toplot_acute$Seq_name))
```


```{r}
ggplot() +
  geom_rect(data=rects, aes(ymin=ystart, ymax=yend, xmin=-Inf,
                      xmax=Inf, fill=col), alpha =0.5) +
  geom_vline(xintercept=0, size=0.25, alpha=0.7) +
  geom_point(data=tiny_toplot_acute, aes(x=LFC, y=Seq_name, color=Color), size=3) +
  scale_color_manual(values = c("blue", "grey", "red")) +
  scale_fill_manual(values=c("grey", "white")) +
  scale_y_discrete(limits=rev,
                   drop=FALSE,
                   #labels=addline_format(tiny_toplot[tiny_toplot$Group == "Acute","Seq_name"]),
                   ) +
  theme_light() +
  facet_wrap(~ Time) +
  theme(axis.text.y = element_text(family = "Courier", size=12),
        legend.position="none",
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(size = 0.3, linetype = "dashed"),
        ) + xlim(-10,13) + ylab("")

ggsave("/path/to/miRNA_and_tiny_results/tiny_acute_plot.pdf", width=8, height=9)
```

```{r}
tiny_toplot_mild <- tiny_toplot_acute <- tiny_toplot[tiny_toplot$Group == "Moderate",]
tiny_toplot_mild$Seq_name <- factor(tiny_toplot_mild$Seq_name, levels=unique(tiny_toplot_mild$Seq_name))
```

```{r}
#rects_2 <- data.frame(ystart=c(0, 2.5), yend=c(2.5,5), col=c("grey", "white"))
```

```{r}
ggplot() +
  # geom_rect(data=rects_2, aes(ymin=ystart, ymax=yend, xmin=-Inf,
  #                     xmax=Inf, fill=col), alpha =0.5) +
  geom_vline(xintercept=0, size=0.25, alpha=0.7) +
  geom_point(data=tiny_toplot_mild, aes(x=LFC, y=Seq_name, color=Color), size=3) +
  scale_color_manual(values = c("grey", "red")) +
  #scale_fill_manual(values=c("white", "grey")) +
  scale_y_discrete(limits=rev,
                   drop=FALSE,
                   #labels=addline_format(tiny_toplot[tiny_toplot$Group == "Acute","Seq_name"])
                   ) +
  theme_light() +
  facet_wrap(~ Time) +
  theme(axis.text.y = element_text(family = "Courier", size=12),
        legend.position="none",
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(size = 0.3, linetype = "dashed"),
        ) + xlim(-10,13) + ylab("")

ggsave("/path/to/miRNA_and_tiny_results/tiny_mild_plot.pdf", width=8, height=2)
```

